---
title: "lab_02"
author: "Shreya Vuttaluru"
date: "9/9/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## About this lab

To complete this lab, you need to:
* run existing code as directed (look for **Task**).
* modify existing code as directed (look for **Task**).
* write code in empty codeblocks provided to answer questions included (look for **Q**).
* write out the answer in the form of a complete sentence in the space given (look for **A**).

When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

## Load libraries and establish settings

**Task** Run the codeblock below to load the Tidyverse and Lubridate. To run the codeblock, hit the little green play button in the upper right corner of the gray area (the codeblock) below.

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse and lubridate
library(tidyverse)
library(lubridate)
```

## Load Data

Let's keep working with West Virginia loans contained in the PPP loans data.  

**Task** Run the codeblock below to load the data.

```{r}
west_virginia_ppp <- read_rds("lab_02.rds")
```

## Examine Data

## Answer questions

**Q1** Create a new column that divides the loan amounts into multiple options: under_10k, 10k_50k, 50k_100k, 100k_500k, 500k_1m, 1m_plus. How many loans are in each group? You will need to use `case_when` to make this work and there's an example of this in the book chapter on mutate.
**A1** [There were 9,310 loans under 10k dollars, 15,372 loans between 10k dollars and 50k dollars, 3,978 loans between 50k dollars and 100k dollars, 4,152 loans between 100k dollars and 500k dollars, 540 loans between 500k dollars and 1 million dollars, and 356 loans that were over 1 million dollars.]

```{r}
# Write the code to produce your answer here
wv_loan_divide <- west_virginia_ppp %>%
  mutate(
    loan_bracket = case_when(
        amount < 10000 ~ "under_10k",
        amount >= 10000 & amount < 50000 ~ "10k_50k",
        amount >= 50000 & amount < 100000 ~ "50k_100k",
        amount >= 100000 & amount < 500000 ~ "100k_500k",
        amount >= 500000 & amount < 1000000 ~ "500k_1m",
        amount > 1000000 ~ "1m_plus"
      )

  ) %>%
  group_by(loan_bracket)  %>%
  summarise(total = n())

```

**Q2** Which servicing lenders have the most loans with undisbursed amounts greater than 0?
**A2** [Harvest Small Business Finance, LLC, had 139 loans with undisbursed amounts greater than 0, more than any other servicing lender.]

```{r}
# Write the code to produce your answer here

west_virginia_ppp %>% 
  mutate(
    undisb_morethan_zero = if_else(
      undisbursed_amount > 0, "Yes", "No"
      )
  ) %>% filter(undisb_morethan_zero == "Yes") %>%
        group_by(servicing_lender_name) %>% 
        summarize(
          count_undibloans = n()
        ) %>%
        arrange(desc(count_undibloans))
      

```

**Q3** How would you describe the top 10 list of those lenders? What do they have in common or not? (You may want to look up some of the top ones.)
**A3** [I noticed that many of the lenders in the top 10 explicitly advertise that they help with PPP loans. Most of them deal with mortgages and finances, but a few of them are also banks. Some of them are also aimed at helping small businesses in particular. Overall, the lenders are situated in different areas of the U.S., and there doesn't seem to be a correlation with location.]

**Q4** Which month in 2021 had the most loans approved?
**A4** [March 2021 had the most loans approved, with 4,404 loans.]

```{r}
# Write the code to produce your answer here
west_virginia_ppp %>%
  mutate(Month = floor_date(date_approved, "month")) %>%
  group_by(Month) %>%
  summarise(total = n()) %>%
  arrange(desc(total))
```
