---
title: "lab_08"
author: "derek willis"
date: "8/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

* A Census API key

## Load libraries and establish settings

You'll need to load two packages for this: the tidyverse and tidycensus.

**Task** load these two packages

```{r}
# Turn off scientific notation
options(scipen=999)
```

## Setup Census API

You'll need your Census API Key:

**Task** set your API Key

```{r echo=FALSE}

library(tidyverse)
library(tidycensus)

#api 
census_api_key("82c33dd2db316b78367d3476bcb90d2abe65a69a", install=TRUE, overwrite=TRUE)

```

## Load and view ACS variables

You'll also need to load the list of American Community Survey variables from the 2019 5-year estimates so you can identify the codes you need:

**Task** load the variables from the 2019 5-year ACS and save them to a dataframe

```{r}
# Write code here to load the ACS variables

acs5 <- load_variables(2019, "acs5", cache = TRUE)

```

## Load PPP data

You'll also need to load an RDS file of PPP loan applications that were removed from the original data by lenders in July.

**Task** read the RDS file and save it to a dataframe

```{r}
# Write code here to read the RDS file

july_ppp <- read_rds("data/lab_08.rds")

```

## Answer questions

**Q1** What is the Maryland jurisdiction with the lowest median age, according to the ACS 2019 5-year estimates? You will need to find the variable for median age first.
**A1**Baltimore City, Maryland. 

```{r}

med_age <- get_acs(geography = "county", state="MD", variables = "B01002_001", year = 2019) 

med_age %>%
  arrange(estimate)
  


```

**Q2** Which Maryland counties have a median age of at least 50?
**A2** Talbot and Worcester County. 

```{r}
med_age %>%
  filter(estimate >= 50)

```

**Q3** Which Maryland counties have the highest percentage of 5-17 year-olds that speak Spanish at home? You'll need one dataframe with county-level data on the number of 5-17 year-olds that speak Spanish at home and a second dataframe with the total population of 5-17 year-olds by county. For each dataframe, rename the `estimate` column to a meaningful name you've given the ACS variable, remove the `variable` columns in each using select and then join them based on GEOID and NAME. Then you can add a column to calculate the percentage of the total 5-17 population that speaks Spanish at home.
**A3** Prince George's and Montgomery counties have the most percentage of 5-17 year olds that speak spanish at home. 

```{r}

## speak spanish 5 - 17 ## B16007_004
spanish_5_17 <- get_acs(geography = "county", state="MD", variables = "B16007_004", year = 2019) 

all_5_17 <- get_acs(geography = "county", state="MD", variables = "B16007_002", year = 2019) 

#renaming 

speak_5_17 <- spanish_5_17 %>%
  rename(speak_5_17=estimate)

all_children <- all_5_17 %>%
  rename(all_children = estimate)

## joining

all_children%>%
  inner_join(speak_5_17, by=c("GEOID", "NAME")) %>%
  select(all_children, speak_5_17, moe.x, moe.y, GEOID, NAME) %>%
  mutate(pct_spanish = (speak_5_17/all_children ) * 100) %>%
  arrange(desc(pct_spanish))


```

**Q4** Given the counties with the largest percentage of 5-17 year-olds that speak Spanish at home, what stands out to you about the data? How would you describe the top 5 counties? How would you describe Garrett County's numbers?
**A4** All of the top 5 counties are fairly diverse, which makes the likelihood of having more spanish speakers increase. Prince George's and Montgomery Counties might be outliers because they are so diverse, which Caroline, Frederick, and Talbot are only marginally higher than the other counties. Garrett County just probably does not have many people of color or a robust school language program, which is why its numbers are so low. 

**Q5** Which 5-digit zip code had the most removed PPP applications, and what is the percentage of Black population in that zip code according to the ACS? To answer this question, you will need to create separate dataframes with ACS data on total population by zip code and Black population by zip code. In each dataframe, rename the `estimate` column and remove the `variable` column. You'll need to combine the two dataframes in order to calculate the percentage, and then you can combine it with the PPP data. Remember: the PPP data may not have only 5-digit zip codes!
**A5** This zipcode is 60620, and is in Chicago. The population is 96.46 percent Black. 

```{r}

# zip codes
## this gives you the top zip codes with removed loans
removed_july_ppp <- july_ppp %>%
  mutate(zip5 = str_sub(zip, 1, 5)) %>%
  group_by(zip5) %>%
  summarize(
    total = n()
  ) %>%
  arrange(desc(total))

# total pop by zip code #	
total_pop <- get_acs(geography = "zcta", variables = "C02003_002", year = 2019) 

# black population by zip code ##	C02003_004
black_pop <- get_acs(geography = "zcta", variables = "C02003_004", year = 2019) 

# joinin

black_pop %>%
  inner_join(total_pop, by=c("GEOID", "NAME")) %>%
  rename(total_pop = estimate.y, black_pop = estimate.x)  %>%
  select(-variable.x, -variable.y, -moe.x, -moe.y) %>%
  mutate(pct_black_pop = (black_pop/total_pop) * 100) %>%
  arrange(desc(pct_black_pop)) %>%
  filter(GEOID == "60620")
  

```

**Q6** How would you describe to an editor the top 10 zip codes with the most removed application in terms of Black population? Where are they located, and what do they have in common? Are there any zip codes that seem different, and if so why?
**A6** These zipcodes had a large Black population, and are located in Chicago, or in Illinois, around the same area. 


**Q7** What other columns from the PPP data could be useful in reporting on what happened to this removed loans?
**A7** The date approved, loan status, location and lender columns could be most useful in reporting on what happened to the removed loans, in order to give us more information about the loan removal and find patterns among removed loans. 
