---
title: "Reverse Engineering Project"
author: "Shreya, Kylie, and Michael"
date: "Date here"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, [Over a Dozen Black and Latino Men Accused a Cop of Humiliating, Invasive Strip Searches. The NYPD Kept Promoting Him.](https://www.propublica.org/article/over-a-dozen-black-and-latino-men-accused-a-cop-of-humiliating-invasive-strip-searches-the-nypd-kept-promoting-him)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}

library(tidyverse)
library(lubridate)

#install.packages("readxl")

library(readxl)

```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

* id
* two
* three 

```{r}
# Load required data

# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

# Clean required data and prepare for analysis if needed. 

allegations <- read_csv("data/CCRB-Complaint-Data_202007271729/allegations_202007271729.csv") %>%
  mutate(
    clean_disposition = case_when(
    str_detect(board_disposition, "^S") ~ "Substantiated",
    TRUE ~ board_disposition
  )
  ) %>% 
  #new column that identifies officers that are currently a higher rank than captain
  mutate(
    rank_above_captain = if_else(
      rank_now == 'Deputy Inspector' | 
      rank_now == 'Inspector' | 
      rank_now == 'Deputy Chief' | 
      rank_now == 'Assistant Chief' | 
      rank_now == 'Bureau Chief' | 
      rank_now == 'Police Comissioner' | 
      rank_now == 'Chiefs and other ranks', "Yes", "No")
  )


  
  allegation_description <- read_excel("data/CCRB-Complaint-Data_202007271729/ccrb_data_layout_table.xlsx")




```

## Data Dictionary
Our data dictionary describes what each variable (column name) means in the cleaned_allegations dataframe and gives the number of variable types for each variables. The only variables outlined in the dictionary are the ones used in order to reverse engineer the article.

The following structure is used for our data dictionary:
variable_name: Description of variable (# of variable types)

unique_mos_id: The unique identification number given to each Member of Service in the NYPD (3996)
first_name: The first name of the Member of Service in the NYPD who has received a complaint, limited to 10 characters (1217)
last_name: The last name of the Member of Service in the NYPD who has received a complaint, limited to 10 characters (2835)
rank_now: The current rank of the Member of Service in the NYPD who has received a complaint (8)
mos_ethnicity: The ethnicity of the Member of Service in the NYPD who has received a complaint (5)
allegation: A brief description of the allegation made against a Member of Service in the NYPD, not cleaned (115)
board_disposition: The CCRB's decision regarding a complaint against a Member of Service in the NYPD, not cleaned (11)
cleaned_board_disposition: The CCRB's decision regarding a complaint against a Member of Service in the NYPD, cleaned for less variable types (3)
rank_above_captain: The current rank of the Member of Service in the NYPD who has received a complaint cleaned into above captain or not (2)
grouped_allegation: A description of the allegation made against a Member of Service in the NYPD, cleaned for less variable types (8)

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.


## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: [ “A dozen of those high-ranking officers have three or more complaints with substantiated allegations; 10 are white. Five have reached the rank of deputy inspector, the lowest promoted at the commissioner’s discretion; all are white.”]
* **Analysis summary**: [My analysis did not confirm the results from the article, but I believe this is due to the fact that the data has been updated since. I found that there were 25 high-ranking officers with three or more complaints, not a dozen, and 17 were White, not 10. Also, there were 11 high-ranking officers with three or more complaints that are now a deputy inspector, and they are not all White. In regard to format, I narrowed down the data to the appropriate requirements, then grouped and summarized to obtain the number of total complaints for each officer. Finally, I filtered for officers with more than two complaints.]

```{r}
# Put code to reverse engineer sentence here


#glimpsing to find rows
glimpse(allegations)


# filtering !
## text says ranks above captain, so not including those. 

## below is high-ranking officers with more than 3 complaints:
highrankallegations <- allegations %>%
filter(rank_above_captain == "Yes") %>%
filter(clean_disposition == "Substantiated") %>%
    group_by(unique_mos_id, first_name, last_name, rank_now, mos_ethnicity) %>%
  summarize(
    total_allegations = n()
  ) %>%
    ungroup() %>%
    filter(total_allegations > 2) %>%
    arrange(desc(total_allegations)) %>%
    summarize(
          total_high_ranking = n()
        )

highrankallegations

## filter by race (# of white officers)
allegations %>%
  filter(rank_above_captain == "Yes") %>%
  filter(clean_disposition == "Substantiated") %>%
  filter(mos_ethnicity == "White") %>%
  group_by(unique_mos_id, first_name, last_name, rank_now, mos_ethnicity) %>%
    summarize(
      total_allegations = n()
    ) %>%
    ungroup() %>%  
    filter(total_allegations > 2) %>%
      arrange(desc(total_allegations)) %>%
      summarize(
        total_high_ranking_white = n()
       )

#above shows 17 white officers. 

## those that reached deputy inspector 
allegations %>%
  filter(rank_now == "Deputy Inspector") %>%
  filter(clean_disposition == "Substantiated") %>%
  group_by(unique_mos_id, first_name, last_name, rank_now, mos_ethnicity) %>%
    summarize(
      total_allegations = n()
    ) %>%
      ungroup() %>%
      filter(total_allegations > 2) %>%
      arrange(desc(total_allegations)) %>%
      summarize(
        total_deputy_inspector = n()
      )



# Display results of code below this codeblock 

```

### Sentence 2

* **Sentence text**: [“Eighty-six of the roughly 420 officers in the department who currently hold a rank above captain — running precincts and other large commands and overseeing hundreds of officers — have tallied at least one misconduct allegation that was substantiated by the CCRB, meaning that investigators amassed enough evidence of offenses, ranging from bad language to pistol whippings, to say that they happened and broke patrol guidelines. The most common involved improper property entries and searches.”]
* **Analysis summary**: [In this section of code, we began by looking into the allegations by rank, filtering for ranks above captain and substantiated claims from the new cleaned columns. Then, I grouped by officer and displayed a count, allowing us to reach the conclusion that 81 of around 420 officers in the department who hold a rank above captain have had at least one misconduct allegation that was substantiated by the CCRB. The most common misconduct allegation was in fact imporoper proterty entries and searches, but this could have been different if the allegations were cleaned in a different format.]

```{r}
# Put code to reverse engineer sentence here

#what are the ranks
glimpse(allegations$rank_now)


# finding the sustained allegations 
allegations %>%
  filter(rank_above_captain == "Yes") %>%
  filter(clean_disposition == "Substantiated") %>%
  group_by(unique_mos_id, first_name, last_name, rank_now, mos_ethnicity) %>%
  summarize(
    total_allegations = n()
  )  %>%
  ungroup() %>%
  summarize(
        total_high_ranking_substantiated = n()
      )

#most common misconduct allegation
allegations %>%
  filter(rank_above_captain == "Yes") %>%
  filter(clean_disposition == "Substantiated") %>%
  group_by(allegation) %>%
  summarize(
    count = n()
  ) %>%
  arrange(desc(count))


# Display results of code below this codeblock

```

### Sentence 3

* **Sentence text**: ["Of at least 77 allegations made against him in 26 separate CCRB complaints, 29 were unsubstantiated; five were “unfounded,” meaning investigators concluded the incident never took place; and 27 were “exonerated,” meaning the conduct fell within patrol guidelines."]
* **Analysis summary**: [In order to confirm/deny this statement, we needed to see how many claims Christopher McCormack had, so we filtered using his ID (18589) to develop a new data frame called mccormack_allegations. Using this dataframe, we filtered for unsubstantiated claims first, returning a result of 29. Then we tried for exonerated claims, which returned 27. Unfounded claims were not present in the dataset provided, so we cannot evaluate that point further; however, we confimred the other results from the sentence.
]

```{r}
# Put code to reverse engineer sentence here
#make new dataset that shows only Christopher McCormack's allegations
#Christopher McCormack's unique_mos_id is 18589
chrisallegations <- allegations %>%
  filter(unique_mos_id == 18589)


# # of unsubstantiated allegations for chris
chrisallegations %>% 
  filter(clean_disposition == "Unsubstantiated") %>%
  group_by(unique_mos_id) %>%
  summarize(
    unsubstantiated_count = n()
  )

## 29 !!

## unfounded claims were not entered into the dataset 


## below codeblock shows the types of allegations (including exonerated)
allegations %>% 
  filter(last_name == "Mccormack",
         first_name == "Christophe") %>% 
  group_by(clean_disposition) %>% 
   summarize(
    count = n())  %>%
    arrange(desc(count)) 
  


# Display results of code below this codeblock

```

### Sentence 4

* **Sentence text**: [But many of the officers also said that it’s possible to be an effective cop without so many complaints. An overwhelming majority of the city’s 36,000 officers do not have a single substantiated complaint on their records; only 1 in 9 does.]
* **Analysis summary**: [For this sentence, we began by creating an entirely new dataset, called officers_with_complaints, that would return the total number of officers with at least one substantiated complaint.  First, we took the original dataframe and  filtered for substantiated allegations in our new cleaned_board_disposition column, then used distinct to return only the 'unique_mos_id', 'first_name', 'last_name' and 'rank_now' columns. After this, we used a summarize function to get the total number of officers with at least one substantiated complaint, which we then divided by the total number of officers (36,000). Our result was .1111, which is equal to 1 in 9, confirming the results from the story.]

```{r}
# Put code to reverse engineer sentence here

#make new dataset that gives number of officers with at least one substantiated complaints
officers_with_complaints <- allegations %>%
  filter(clean_disposition == "Substantiated") %>%
  distinct(unique_mos_id, first_name, last_name, rank_now)
#number of officers with at least one substantiated complaint/total number of officers
officers_with_complaints %>%
    summarize(
      total = n()/36000
    ) 
#1 in 9 = .1111

# Display results of code below this codeblock

officers_with_complaints 

```

### Sentence 5

* **Sentence text**: [There isn't really a sentence from the story, but I'd like you to answer this question: The story focused on Christopher McCormack. Use data analysis to justify that decision. Why does he stand out as newsworthy, from all of the people they could have selected? Why, do you think, did the reporters single him out for an in-depth story. Write code to support your argument.]
* **Analysis summary**: [ In our last section, we used code to show why Christopher McCormack stands out as newsworthy. In order to do this, we used our dataframe to filter for all officers above the rank of captain, or in other words, high-ranking officers. Then we grouped by officer to count the total number of allegations for each, in which Christopher McCormack is the second to appear of the sorted list. Not only does he have the second most number of allegations, but he is higher ranking than the only person with more than him, making him more newsworthy as the question as to why he was continuously promoted becomes apparent.

```{r}
# Put code to reverse engineer sentence here
cleaned_allegations %>%
  filter(rank_above_captain == "Yes") %>%
  group_by(unique_mos_id, first_name, last_name, rank_now, mos_ethnicity) %>%
  summarize(
    total_allegations = n()
  ) %>%
  arrange(desc(total_allegations))
```

-30-