---
title: "Data Analysis Project"
author: "michael purdie, kylie rau, shreya vuttaluru"
date: "11/14/21"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are working with the [Washington Post police shooting database](https://github.com/washingtonpost/data-police-shootings) and the [Washington Post homicides database.](https://github.com/washingtonpost/data-homicides)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
#essential libraries
library(tidyverse)
library(janitor)
library(lubridate)
#additional libraries 
library(sf)
library(tigris)
library(tidycensus)

install.packages("usmap")
library(usmap)

install.packages("censusxy")
library(censusxy)

```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

## Quick Analysis based on loaded data: in looking at the police shootings data, everything looks relatively clean, so we don't foresee having to do any cleaning. We're choosing to focus on the geography of police shootings, so we need to use census data in order to figure out where exactly these police shootings are happening, and if there are certain neighborhoods (ex: majority-minority or low income) that are more affected by police shootings. We'll likely need to form some kind of relationship between the lat/long coordinates in this dataset and data from census tracts. 

```{r}
# Load required data
police_shootings <- read_csv("data/data-police-shootings-master/fatal-police-shootings-data.csv")

## census api key
census_api_key("82c33dd2db316b78367d3476bcb90d2abe65a69a", install = "TRUE", overwrite = TRUE)

variables <- load_variables("acs5", year=2019)

```

## Working with the data 

## Deliverable 3 Summary: This week, we spent most of our time figuring out how to reverse geocode the coordinates given by the Washington Post so that we can eventually retrieve characteristics about the neighborhoods (census tracts) that the shootings occur in. Right now, we don’t have any particular findings, but now that we have geoid as its own separate column within our data, we’ll be able to call in census data via tidycensus and other packages and potentially join our tables in order to better narrow down the characteristics of neighborhoods that police shootings occur in more often. From there, we’ll have to do a little bit of sorting and filtering to better understand the data. 

``` {r}
## trying to make a map to generally understand where the shootings are
## getting some county data
counties <- counties() 
usa_counties <- counties %>%
  filter (STATEFP < "57")

## trying to fuck around with census xy

police_shootings <- read_csv("data/data-police-shootings-master/fatal-police-shootings-data.csv") %>%
    filter(!is.na(longitude)) 

## ok so it seems like we can only do one lat long at a time so... for loop? idk lets try

## first create empty data frame

police_shootings_sf <- tibble()

for (row_number in 1:nrow(police_shootings)) {
  
  #this makes a dataframe for each
  row_df<- police_shootings %>%
  slice(row_number)
  
  #store lat and long values
  longitude <- row_df$longitude
  latitude <- row_df$latitude
  census_results <- cxy_geography(longitude, latitude) %>%
    select(Census.Tracts.GEOID) %>%
    clean_names()
 
   row_df <- row_df %>%
     bind_cols(census_results) 
   
  
  #inding some rows
   police_shootings_sf <- police_shootings_sf %>%
    bind_rows(row_df) 
   
   print(paste0("finished ", row_number, " ", Sys.time()))
   
   if (row_number%%500 == 0) {
     filepath <- paste0("data/geocoded_results_", row_number, ".rds")
     write_rds(police_shootings_sf, filepath)
     police_shootings_sf <- as_tibble()
     
   }
     

}

write_rds(police_shootings_sf, "data/geocoded_results.rds")


## testingggggg
read_rds("data/geocoded_results_2500.rds")

####### START TEST

## testing new data with 2019 geocodes
    data_test <- read_rds("data/new_geocoded_results_500.rds")
    data_test2 <- read_rds("data/new_geocoded_results_4000.rds")
   
     test_census <- get_acs(geography = "tract",
                       variables = c(population = "B01001_001", 
                                     median.gross.rent = "B25064_001",
                                     median.household.income = "B19013_001",
                                     rent.burden = "B25071_001",
                                     white = "B03002_003",
                                     af.am = "B03002_004",
                                     hispanic = "B03002_012",
                                     am.ind = "B03002_005",
                                     asian = "B03002_006",
                                     nh.pi = "B03002_007",
                                     multiple = "B03002_009",
                                     poverty_level ="B06012_002",
                                     other = "B03002_008"),
                       state = state_list, 
                       year = 2019)


    test_census <- pivot_wider(test_census, names_from = variable, names_sep = ".", values_from = c(estimate, moe)) 

    #rename column
    test_census <- rename(test_census, census_tracts_geoid = GEOID)

    #join
    test_w_acs <- data_test %>%
    left_join(test_census, by=c("census_tracts_geoid")

    ## nice!
    
######### END TEST !!!!
        
######### Reading in data 

data1 <- read_rds("data/new_geocoded_results_500.rds")
data2 <- read_rds("data/new_geocoded_results_1000.rds")
data3 <- read_rds("data/new_geocoded_results_1500.rds")
data4 <- read_rds("data/new_geocoded_results_2000.rds")
data5 <- read_rds("data/new_geocoded_results_2500.rds")
data6 <- read_rds("data/new_geocoded_results_3000.rds")
data7 <- read_rds("data/new_geocoded_results_3500.rds")
data8 <- read_rds("data/new_geocoded_results_4000.rds")
data9 <- read_rds("data/new_geocoded_results_4500.rds")
data10 <- read_rds("data/new_geocoded_results_5000.rds")
data11 <- read_rds("data/new_geocoded_results_5500.rds")
data12 <- read_rds("data/new_geocoded_results_6000.rds")
data13 <- read_rds("data/new_geocoded_results_6410.rds")

## binding rows (making a full dataset)
shootings_w_geoid <- bind_rows(data1, data2, data3, data4, data5, data6, data7, data8, data9, data10, data11, data12, data13, .id = NULL)

## antijoining to figure out unmatched rows
unmatched <- anti_join(police_shootings, shootings_w_geoid, by=c("id"))
## only one match which is Gary Brown 

## testing cxy_vintages
testing <- cxy_geography(lon = -122.422,
              lat = 37.763, 
              benchmark = "Public_AR_Current",
              vintage = "ACS2019_Current")


## so we have 1 less than the dataset, but it's Gary Brown, who was a bum geocode. 

### joining census data

state_list <- fips_codes %>%
  distinct(state) %>%
  head(51) %>%
  as_vector()

census_data <- get_acs(geography = "tract",
                       variables = c(population = "B01001_001", 
                                     median.gross.rent = "B25064_001",
                                     median.household.income = "B19013_001",
                                     rent.burden = "B25071_001",
                                     white = "B03002_003",
                                     af.am = "B03002_004",
                                     hispanic = "B03002_012",
                                     am.ind = "B03002_005",
                                     asian = "B03002_006",
                                     nh.pi = "B03002_007",
                                     multiple = "B03002_009",
                                     poverty_level ="B06012_002",
                                     other = "B03002_008"),
                       state = state_list, 
                       year = 2019)


#PIVOT
census_data <- pivot_wider(census_data, names_from = variable, names_sep = ".", values_from = c(estimate, moe)) 

#rename column
census_data <- rename(census_data, census_tracts_geoid = GEOID)

#join
shootings_w_acs <- shootings_w_geoid %>%
  left_join(census_data, by=c("census_tracts_geoid"))

## making an .rds dataset so it's easy for my friends to download :)

write_rds(shootings_w_acs, "data/new_shootings_w_acs.rds")

### STARTING ANALYSIS HERE: 

## first looking at data for highest percentage african american tracts: 

black_neighborhood_shootings <- shootings_w_acs %>%
  mutate(
    pct_black = (estimate.af.am/estimate.population) * 100)  %>%
    filter(pct_black > 50) %>%
  select(name, age, date, race, city, state, NAME, pct_black) %>%
  arrange(desc(pct_black))

## 682 shootings in neighborhoods that are majority black in a dataset of 6410

(682/ 6410) * 100
## interesting --> 10.63 percent of the total shootings were in majority black neighborhoods. 

## how many of those were black people? 
black_neighborhood_shootings %>%
  filter(race == "B") %>%
  summarize (
    count = n() 
  ) %>%
  as_tibble()

## 420 of those are black 
(509 / 682) * 100
## which is 74.63 percent of the shootings in Black neighborhoods

## where were the most shootings? 
    #tract level
black_neighborhood_shootings %>%
  group_by(NAME) %>%
   summarize(
   count = n() 
  ) %>%
  arrange(desc(count))
   ## Duval County, Florida --> Jacksonville

    #city level
black_neighborhood_shootings %>%
  group_by(city) %>%
   summarize(
   count = n() 
  ) %>%
  arrange(desc(count))

  ## on the city level, most shootings are in Chicago, and many of them were in major cities 

## racial breakdown of all of the shootings
shootings_w_acs %>%
   group_by(race) %>%
   summarize(
   count = n() 
  ) %>%
  arrange(desc(count))
## overall, white people made up more of the shootings. 

## breakdown in majority black neighborhoods
black_neighborhood_shootings %>%
  group_by(race) %>%
   summarize(
   count = n() 
  ) %>%
  arrange(desc(count))


## income breakdown !! 
    ## Pew Research Center defines low income households as those earning below $48,500 
    
## general breakdown first:
low_income_shootings <- shootings_w_acs %>%
  filter(estimate.median.household.income < 48500) %>% 
   summarize(
   count = n() 
  ) 
## 2924 shootings out of 6409 were in low-income neighborhoods 
(2924/6409) * 100
## that's 45 percent !! 

## where are the specific places:
low_income_shootings %>%
  group_by(NAME) %>%
   summarize(
   count = n() 
  ) %>%
  arrange(desc(count))


## white people shot in majority white neighborhoods
white_neighborhood_shootings_white <- shootings_w_acs %>%
filter(race == "W", estimate.white > estimate.af.am + estimate.am.ind + estimate.asian + estimate.nh.pi + estimate.other)

## white people shot in majority white neighborhoods
white_neighborhood_shootings_black <- shootings_w_acs %>%
filter(race == "B", estimate.white > estimate.af.am + estimate.am.ind + estimate.asian + estimate.nh.pi + estimate.other)


## total people shot in majority white neighborhoods
majority_white_neighborhood_shootings <- shootings_w_acs %>%
filter(estimate.white > estimate.af.am + estimate.am.ind + estimate.asian + estimate.nh.pi + estimate.other)

1887/3599 = 0.5243123 ##will be different with correct data
= 52.5% of people shot in majority white neighborhoods are white.

527/3599 = 0.1464296 ##will be different with correct data
= 14.6% of people shot in majority white neighborhoods are black.


## breakdown in majority black neighborhoods
black_neighborhood_shootings %>%
  group_by(race) %>%
   summarize(
   count = n() 
  ) %>%
  arrange(desc(count))
  ##black people made up the majority
  
## white people shot in majority white neighborhoods
white_neighborhood_shootings_white <- shootings_w_acs %>%
filter(race == "W", estimate.white > estimate.af.am + estimate.am.ind + estimate.asian + estimate.nh.pi + estimate.other)
white_neighborhood_shootings_white
  ##2573 people
## black people shot in majority white neighborhoods
white_neighborhood_shootings_black <- shootings_w_acs %>%
filter(race == "B", estimate.white > estimate.af.am + estimate.am.ind + estimate.asian + estimate.nh.pi + estimate.other)
white_neighborhood_shootings_black
  ##709 people
## total people shot in majority white neighborhoods
majority_white_neighborhood_shootings <- shootings_w_acs %>%
filter(estimate.white > estimate.af.am + estimate.am.ind + estimate.asian + estimate.nh.pi + estimate.other)
majority_white_neighborhood_shootings
  ##4856 people
2573/4856 = 0.52986*100 
    = 52.986% of people shot in majority white neighborhoods are white. 
709/4856 = 0.1460049*100
    = 14.6% of people shot in majority white neighborhoods are black. 
## black people shot in majority black neighborhoods
black_neighborhood_shootings_black <- shootings_w_acs %>%
filter(race == "B", estimate.af.am > estimate.white + estimate.am.ind + estimate.asian + estimate.nh.pi + estimate.other)
black_neighborhood_shootings_black
  ##663 people
## white people shot in majority black neighborhoods
black_neighborhood_shootings_white <- shootings_w_acs %>%
filter(race == "W", estimate.af.am > estimate.white + estimate.am.ind + estimate.asian + estimate.nh.pi + estimate.other)
black_neighborhood_shootings_white
  ##144 people
## total people shot in majority black neighborhoods
majority_black_neighborhood_shootings <- shootings_w_acs %>%
filter(estimate.af.am > estimate.white + estimate.am.ind + estimate.asian + estimate.nh.pi + estimate.other)
majority_black_neighborhood_shootings
  ##1029 people
144/1029 = 0.1399417*100 
    = 13.4% of people shot in majority black neighborhoods are white. 
663/1029 = 0.6443149*100
    = 64.4% of people shot in majority black neighborhoods are black. 
```


### Income Analysis

``` {r}
## income breakdown !! 
    ## Pew Research Center defines low-income households as those earning below $48,500 
    ## Pew Research Center defines high-income households as those earning above $145,500
    
## general low-income breakdown first:
low_income_shootings <- new_shootings_w_acs %>%
  filter(estimate.median.household.income < 48500)
low_income_shootings %>% 
   summarize(
   count = n() 
  ) 
  ## 2924 shootings out of 6409 were in low-income neighborhoods 
(2924/6409) * 100
  = 45.6% of shootings were in low-income neighborhoods.
## where are the specific places:
low_income_shootings %>%
  group_by(NAME) %>%
   summarize(
   count = n() 
  ) %>%
  arrange(desc(count))
  ##none of these seem to be significant
## white people shot in majority low-income neighborhoods
low_income_neighborhood_shootings_white <- new_shootings_w_acs %>%
filter(race == "W", estimate.median.household.income < 48500)
low_income_neighborhood_shootings_white
  ##1108 people
## black people shot in majority low-income neighborhoods
low_income_neighborhood_shootings_black <- new_shootings_w_acs %>%
filter(race == "B", estimate.median.household.income < 48500)
low_income_neighborhood_shootings_black
  ##865 people
1108/2924 = 0.378933*100 
    = 37.9% of people shot in majority low-income neighborhoods are white. 
865/2924 = 0.2958276*100
    = 29.6% of people shot in majority low-income neighborhoods are black. 
## general high-income breakdown
high_income_shootings <- new_shootings_w_acs %>%
  filter(estimate.median.household.income > 145500)
high_income_shootings %>% 
   summarize(
   count = n() 
  ) 
  ## 59 shootings out of 6409 were in high-income neighborhoods 
(59/6409) * 100
  = 0.92% of shootings were in high-income neighborhoods.
## where are the specific places:
high_income_shootings %>%
  group_by(NAME) %>%
   summarize(
   count = n() 
  ) %>%
  arrange(desc(count))
  ##none of these seem to be significant, except maybe top 3 in Texas?
## white people shot in majority high-income neighborhoods
high_income_neighborhood_shootings_white <- new_shootings_w_acs %>%
filter(race == "W", estimate.median.household.income > 145500)
high_income_neighborhood_shootings_white
  ##33 people
## black people shot in majority high-income neighborhoods
high_income_neighborhood_shootings_black <- new_shootings_w_acs %>%
filter(race == "B", estimate.median.household.income > 145500)
high_income_neighborhood_shootings_black
  ##13 people
33/59 = 0.378933*100 
    = 55.9% of people shot in majority high-income neighborhoods are white. 
13/59 = 0.2958276*100
    = 22.0% of people shot in majority high-income neighborhoods are black. 


### some mapping stuff below, is separate and none of it is working right now. 
 
cxy_geography(longitude, latitude)

## mapping all the counties

## this is not really working rn but I am trying lol

ggplot() + 
  geom_sf(police_shootings_geom)
  geom_point(data=police_shootings, aes(x=longitude, y=latitude)) +
  geom_sf(data=usa_counties) + 
  theme_minimal() +
  geometry = geometry

```