---
title: "lab_04"
author: "not derek willis"
date: "8/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## About this lab

To complete this lab, you need to:
* run existing code as directed (look for **Task**).
* modify existing code as directed (look for **Task**).
* write code in empty codeblocks provided to answer questions included (look for **Q**).
* write out the answer in the form of a complete sentence in the space given (look for **A**).

When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

## Load libraries and establish settings

**Task** Run the codeblock below to load the Tidyverse. To run the codeblock, hit the little green play button in the upper right corner of the gray area (the codeblock) below.

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse.
library(tidyverse)
```

## Load Data

Let's keep working with Maryland loan applications contained in the PPP data.

**Task** Read in the lab_04.rds file and assign it to the variable `maryland_ppp`.

```{r}
# add code here to accomplish the task
maryland_ppp <- read_rds("lab_04.rds")

```

To answer the questions below, you'll need to join both the NAICS code data and the ZCTA demographic data from the book chapter. There's a particular set of steps you'll need to perform to make that work properly.

First, let's load the NAICS code CSV file we have in the data folder:

**Task** Read in the naics_codes.csv file and assign it to the variable `naics_codes`.

```{r}
# add code here to accomplish the task

naics_codes <- read_csv("data/naics_codes.csv")

```

Next join the two and then create a 5-character `zip5` column that we can use to join to the ZCTA data:

**Task** Write code to join the maryland_ppp and naics_codes data, assigning the result to a new variable. Then add a new `zip5` column.

```{r}
# add code here to accomplish the task

md_ppp_naics <- maryland_ppp %>% left_join(naics_codes, by="naics_code") %>%
  mutate(zip5 = str_sub(zip, start=1L, end=5L))

glimpse(md_ppp_naics)

  


```

Let's add the Zip Code Tabulation Area dataset. You'll need to make the `ZCTA5N` column a character column, not a numeric one:

**Task** Read in the Zip Code Tabulation Area dataset in the data folder and change the datatype of the `ZCTA5N` column to character.

```{r}
# add code here to accomplish the task

#loading data
md_zcta <- read_csv("data/maryland_zcta.csv")

#making column a character
md_zcta <- md_zcta %>% 
  mutate(across(ZCTA5N, as.character))

md_ppp_naics <- md_ppp_naics %>% 
   mutate(across(zip5, as.character))



```

Now we can join the PPP data to the ZCTA data:

**Task** Join the Maryland PPP data (with NAICS already joined) with the ZCTA data, specifying the join column in both dataframes.

```{r}
# add code here to accomplish the task

md_naics_zcta <- md_ppp_naics %>% left_join(md_zcta, by=c("zip5"="ZCTA5N")) 

```

## Answer questions

**Q1** What are the top 5 industries (using the NAICS title) with the most PPP applications in Maryland?
**A1** [The top 5 industries are Taxi Service (7363 loans), Beauty Salons (7275 loans), Full-Service Restaurants (5490	loans), Offices of Real Estate Agents and Brokers (4443	loans), and All Other Personal Services (3956	loans).]

```{r}
# Write the code to produce your answer here

md_naics_zcta %>%
  group_by(title) %>%
  summarize(
     count = n())  %>%
  arrange(desc(count))


```

**Q2** What are the top industries (using the NAICS title) with the most PPP applications in Maryland where the loan amount is more than $150,000?
**A2** [Full-Service Restaurants, Offices of Physicians (except Mental Health Specialists), and
Plumbing, Heating, and Air-Conditioning Contractors are among businesses where the loan amount is more than $150,000.]

```{r}
# Write the code to produce your answer here

md_naics_zcta %>%
  filter(amount > 150000) %>%
  group_by(title) %>%
  summarize(
     count = n())  %>%
  arrange(desc(count))

```

**Q3** Which Maryland zip code with at least 20 percent non-Hispanic Asian population had the largest sum of PPP loans? Also show the number (count) of loans.
**A3** [20850. There were 3186 loans.]

```{r}
# Write the code to produce your answer here

md_naics_zcta %>%
  filter(NHA >= 20) %>%
  group_by(zip5) %>%
  summarize(
     count = n())  %>%
  arrange(desc(count))

```

**Q4** Which Maryland zip code had the largest sum of loans where the reported race of the applicant was "Asian", and what is the percentage of non-Hispanic Asian population in that zip code? Also show the number (count) of loans.
**A4** [20850 again. The percentage of non-Hispanic Asian popultion is 23.5 percent. There were 240 loans.]

```{r}
# Write the code to produce your answer here

md_naics_zcta %>%
  filter(race == "Asian") %>%
  group_by(zip5) %>%
  summarize(
     count = n())  %>%
  arrange(desc(count))

md_naics_zcta %>%
  mutate (pct_asian = NHA / POP100 * 100) %>%
  group_by(zip5, NHA, pct_asian)  %>%
  summarize(
     count = n())  %>%
  arrange(desc(count))

```

**Q5** What might explain why 20850 has a significant Asian population (23.5%) and only 240 of the 3,186 loans (7.5%) are to Asian applicants?
**A5** [Many Asian-owned businesses are immigrant-owned, and people may not be comfortable applying for a federal loan if they are not permanent residents of citizens. Also, many applicants did not report race in their applications, which could also lead to the discrepency.]


**Q6** Find any Maryland zip codes that have loans where the PPP data lists the applicants as being Asian but where the zip code does not appear to have Asian population in the ZCTA data? What do you think is going on here?
**A6** [There are 3 zip codes where the applicant is listed as Asian but the zip code does not appear to have Asian population in the ZCTA data. This is probably an applicant error where the applicant provided the incorrect zip code, or it could be an entry/human error.]

```{r}
# Write the code to produce your answer here

md_naics_zcta %>%
  filter(race == "Asian" & 
         NHA == 0) %>%
  group_by(zip5) %>%
  summarize(
     count = n())  %>%
  arrange(desc(count))


```

**Q7** How many industries have exactly one loan in the data (hint: you can filter _after_ using `summarise` on any totals you create). From a newsworthiness standpoint, which of these is most worthy of further exploration, and why?
**A7** [There are 44 industries that have exactly one loan in the data. I think that casinos having only 1 loan in total is pretty suspect, and could be worth looking into. There are 15 casinos in Maryland, and I find it hard to believe that only one of them recieved a loan.]

```{r}
# Write the code to produce your answer here

md_naics_zcta %>%
 group_by(title) %>%
  summarize(
     count = n())  %>%
  arrange(desc(count)) %>%
  filter(count == 1) %>%
    summarize(
     count = n()) 

```
